Using high-resolution variant frequencies to empower clinical genome interpretation

### 2016-10-11 manuscript/manuscript.Rmd
Wording changes

We recognise several limitations of our approach. First, we are limited by our understanding of the prevalence and genetic architecture of the disease in question: this characterisation will vary for different diseases and in different populations, though we illustrate approaches for estimation and extrapolation of parameters. In particular, we must be wary of extrapolating to or from less-well characterised populations that could harbour population-specific founder mutations. Unfortunately, there is no substitute for understanding the genetic architecture of the disease of interest, but when considering a specific variant that has been observed at low frequency in a given reference population, our framework and accompanying web tool provide a transparent approach to define the range of disease architectures that are compatible with the observed data. For example, we may not be able to say definitively that a variant is benign, but we might be confident that it could only be causal if the prevalence of the disease is higher than published estimates, or its penetrance is below a specified value.

Secondly, it is often difficult to obtain accurate penetrance information for reported variants, and it is also difficult to know what degree of penetrance to expect or assume for newly discovered pathogenic variants. Although we would argue that variants with low penetrance have questionable diagnostic utility, our calculator app allows a user to define a range of compatible penetrance for a given allele frequency (see **Supplementary Methods**). Furthermore, methods to estimate variant penetrance from large case and control cohorts have been described previously`r citep("10.1126/scitranslmed.aad5169")`.

>>>>

We recognise several limitations of our approach. First, the approach is limited by our understanding of the prevalence and genetic architecture of the disease in question: this characterisation will vary for different diseases and in different populations, though we illustrate approaches for estimation and extrapolation of parameters. In particular, we must be wary of extrapolating to or from less-well characterised populations that could harbour population-specific founder mutations. Unfortunately, there is no substitute for understanding the genetic architecture of the disease of interest, but when considering a specific variant that has been observed at low frequency in a given reference population, our framework and accompanying web tool provide a transparent approach to define the range of disease architectures that would be compatible with the observed data. For example, we may not be able to say definitively that a variant is benign, but we might be confident that it could only be causal if the prevalence of the disease is higher than published estimates, or if its penetrance is below a specified value.

Secondly, it is often difficult to obtain accurate penetrance information for reported variants, and it is also difficult to know what degree of penetrance to expect or assume for newly discovered pathogenic variants. Although we would argue that variants with low penetrance have questionable diagnostic utility, our calculator app allows a user to define a range of compatible penetrance for a given allele frequency (see **Supplementary Methods**). Furthermore, methods to estimate variant penetrance from large case and control cohorts have been described previously `r citep("10.1126/scitranslmed.aad5169")`.
